---
title: "promotion_assumption_check"
author: "Ke Deng"
date: "2021/4/9"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(knitr)
library(readr)
```

```{r}
curr_employee <- read_csv("/Users/krist/Documents/303 final project/data/black-saber-current-employees.csv")
```

### Verify assumptions

- the cases are independently distributed
- the random effect variables are normally distributed

```{r}
qqnorm(curr_employee$employee_id)
qqline(curr_employee$employee_id, col = "orange", lwd=2)
```

From the Quantile-Quantile plot above, the points follow closely around the qq-line in the middle and deviates from it at the two ends. Given this information, the variable "id" can be assumed to be normally distributed, hence matching the model assumption. 

```{r}
cleaned_data <- curr_employee
cleaned_data <- cleaned_data%>%
  mutate(team, team_num =case_when(
    team == "Client services"~1,
    team == "Data"~2,
    team == "Design Legal and financial"~3,
    team == "Marketing and sales"~4,
    team == "Operations"~5,
    team == "People and talent"~6,
    team == "Software"~7  ))

qqnorm(cleaned_data$team_num)
qqline(cleaned_data$team_num, col = "orange", lwd=2)
```
Creating and checking the y-variable:
```{r}
cleaned_data <- curr_employee
#cleaned_data <- cleaned_data%>%
#  group_by(employee_id) %>%
#  count()
cleaned_data <- cleaned_data%>%
  mutate(role_seniority, role_num =case_when(
    role_seniority == "Entry-level"~1,
    role_seniority == "Junior I"~2,
    role_seniority == "Junior II"~3,
    role_seniority == "Senior I"~4,
    role_seniority == "Senior II"~5,
    role_seniority == "Senior III"~6,
    role_seniority == "Manager"~7,
    role_seniority == "Director"~8,
    role_seniority == "Vice president"~9,
    ))

cleaned_data <- cleaned_data %>%
  group_by(employee_id) %>%
  summarize(num_promotions = max(role_num) - min(role_num))


hist(cleaned_data$num_promotions)

y = cleaned_data$num_promotions
n = length(y)
x=table(y)
k=as.numeric(names(x))
plot(k,log(x)+lfactorial(k))
n*(1-cor(k,log(x)+lfactorial(k))^2)
```
In the histogram -- downward shaping, not linear - Poisson
to further verify, using Poissonness plot and goodness of fit test of a Poisson, the response variable follows a Poisson distribution, and hence all assumptions are matched, we are ready to build the model














